A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-01-24 paper/paper.tex
a rewording

Notice, however, that the recipe obtained so far, $X_{n + 1}: ({S_n})_\kstar(X_n)(X_n)$ for $X_n: S_n$, applies parametricity on the \emph{syntax} of the type of $X_n$. It does not directly yield a characterization of $S_n$ as a function from $n$. Reformulating the recipe as an explicit recursive construction, without requiring an interpretation of the syntax of types, is the main outcome of this work, together with the mechanization and the uniform treatment of augmented semi-simplial and semi-cubical sets by means of the generalization to $\nu$-sets.

@@

Depending on how inequality on natural numbers is defined, different equalities may hold or not definitionally, such as transitivity of inequality or the computational properties of induction over inequality proofs. We do not address these details in this section, see instead~\ref{sec:mech}.

@@

\begin{minted}{coq}
  Fixpoint leR (n m : nat) : SProp :=
  match n, m with
  | O, _ => STrue
  | S n, O => SFalse
  | S n, S m => leR n m
  end.
\end{minted}

Here \SProp is a definitionally proof-irrelevant impredicative universe at the bottom of the universe hierarchy~\cite{gilbert19}. By placing the definition in \SProp, we have definitional equality of all inequality proofs. For the purpose of unification, however, this definition does not go far enough. Consider the unification problems:

\begin{minted}{coq}
  leR_trans ?p leR_refl = ?p
  leR_trans leR_refl ?p = ?p
\end{minted}
where \texttt{leR\_trans} is transitivity, \texttt{leR\_refl} is reflexivity, and \texttt{?p} is an existential variable. These two problems definitionally hold in \SProp, but equalizing them does not solve the existential. For unification to be useful in inferring existentials, we present our first variant of $\leq$, which we dub as the ``Yoneda variant'':

\begin{minted}{coq}
  Definition leY n m :=
    forall p, leR p n -> leR p m.
\end{minted}

@@

\begin{minted}{coq}
  Inductive leI : nat -> nat -> Type :=
  | leI_refl n : n <~ n
  | leI_down {n p} : p.+1 <~ n -> p <~ n
  where "n <~ m" := (leI n m) : nat_scope.
\end{minted}

>>>>

Notice, however, that the recipe obtained so far, $X_{n + 1}: ({S_n})_\kstar(X_n)(X_n)$ for $X_n: S_n$, applies parametricity on the \emph{syntax} of the type of $X_n$. It does not directly yield a characterization of $S_n$ as a function from $n$. Reformulating the recipe as an explicit recursive construction, without requiring an interpretation of the syntax of types, is the main outcome of this work, together with the mechanization and the uniform treatment of augmented semi-simplicial and semi-cubical sets by means of the generalization to $\nu$-sets.

@@

Depending on how inequality on natural numbers is defined, different equalities may hold or not definitionally, such as transitivity of inequality or the computational properties of induction over inequality proofs. Details are given in~\ref{sec:mech}.

@@

\begin{figure}[H]
  \begin{minted}{coq}
  Fixpoint leR (n m : nat) : SProp :=
  match n, m with
  | O, _ => STrue
  | S n, O => SFalse
  | S n, S m => leR n m
  end.
  \end{minted}
\end{figure}

Here $\SProp$ is a definitionally proof-irrelevant impredicative universe at the bottom of the universe hierarchy~\cite{gilbert19}. By placing the definition in \SProp, we have definitional equality of all inequality proofs. For the purpose of unification, however, this definition does not go far enough. Consider the unification problems:

\begin{figure}[H]
  \begin{minted}{coq}
  leR_trans ?p leR_refl = ?p
  leR_trans leR_refl ?p = ?p
  \end{minted}
\end{figure}
where \texttt{leR\_trans} is transitivity, \texttt{leR\_refl} is reflexivity, and \texttt{?p} is an existential variable. These two problems definitionally hold in \SProp, but equalizing them does not solve the existential. For unification to be useful in inferring existentials, we present our first variant of $\leq$, which we dub as the ``Yoneda variant'':

\begin{figure}[H]
  \begin{minted}{coq}
  Definition leY n m :=
    forall p, leR p n -> leR p m.
  \end{minted}
\end{figure}

@@

\begin{figure}[H]
  \begin{minted}{coq}
  Inductive leI : nat -> nat -> Type :=
  | leI_refl n : n <~ n
  | leI_down {n p} : p.+1 <~ n -> p <~ n
  where "n <~ m" := (leI n m) : nat_scope.
  \end{minted}
\end{figure}
